---
title: "FishPrint_workbook"
author: "Jessica Gephart"
date: "6/15/2020"
output:
  word_document: default
  pdf_document: default
---

## Agenda for next meeting

1. Report back from Alon on FCR modeling and FCR-feed composition relationship

2. Yield/harvest data updates
- Patrik going through to update the data in this column
Challenges with data as reported:
- Unclear in a number of studies what the temporal dimension
- Number of harvests per year can vary though
- Some specify it in m3
Decision:
- Set all cages to 0? Amount is very small compared to feeds. 
Focus is on land occupation, but not freshwater or marine area occupation

3. Define groupings
_ Jessica to suggest some incorporating: 
- fed commercial, fed non-commercial and unfed
- intensive, semi-intensitve and extensive
- * Need to update this classification in the data frame - walk through with group for consensus. 
- taxa groups? 

4. Next steps: 
- Standardize units
- Calculate sd's
- Think about biodiversity and how to incorporate
- Think about how to reweight estimates by production/representativeness
- Add in energy N, P, H2O from refineries 
- Handling of byproducts 
- Weighting of feed ingredients:
  - Just use chicken feather meal for animal products
  - Soy is more complicated
- Ocean area required to support 

## To do

1. The biggest gap is harvest and yield, which is needed to get a per unit land FP, as well as a surface area for evaporative losses. 
- Harvest and yield time dimensions will need to be standardized
- Can use data from the Blue Frontiers report
- For places that have multiple uses, we will need to apply an allocation method (consistent with the feeds)

2. Fill in feed ingredient FP and protein content values
 - Alon pulled the terrestrial feed data N and P content --> currently using a straight mean for these
 _ Question for Rob: use of P-eq vs PO4-eq in feed FP data --> check that P-eq would be P release (same for N-eq)
 _ Question for Rob: energy category in feed FP data
 - Question for all: how to weight the footprints of the feed ingredients within a category
 - Top 10 most representative crops (excluding soy b/c that is included separately), check for presence in EcoVent and crop N/P, if not present in both, move to 11th most representative [Patrik will take a stab at this]
 
3. Factors to multiply protein content by to get N and P
 - Patrik can get factors with reference --> done, but may need to add reference
 
4. Whole fish N and P content by species (or use protein content and N and P factors)
 - Alon will talk to Zach about this; also talk to Christina about the modelled values --> done
 
5. Replace placeholder diesel, petrol and natural gas CO2 eq/L values with data in function
 - Rob - diesel already in, others coming

6. What units do we want the FP estimates in?
 - Input from anyone (we just need to pick something at some point to make sure everything is consistent)
 - L, ha, t CO2eq, /kg fish at farm gate

7. Evaporation
 - We have average evaporative loss by country 
 - We may want to consider a multiplication factor for losses from aerated ponds [maybe only include in discussion; would also influence methane emissions from ponds]
 - We need to decide how to construct the distribution (currently just the arithmatic mean, but we could add a weighting)
 - Only apply to freshwater ponds [include exploration of evaporative loss from brackish ponds in SI -- also discuss the dilution water] 

8. Edible portion and per gram protein
  - Present results in terms of LW, but also add in edible portion and per g protein
  - Still need to compile this data
  
9. Add in fish in-fish out metric
 - We can get to this through the same FMFO data that we are using 
 - Return to discussion of broader biodiversity issues [likely just a discussion point -- justification: none of our other categories go all the way to impact]
  
## Code improvements

1. Standardize units throughout
2. Update feed associated FP function to handle fed and non-fed species (or just make all feed ingredient proportions zero for non-fed)
3. Incorporate differences based on system type in each function
4. Add categorical variables we need to add/standardize are: system type, fed/unfed, aerated/non-aearated, intensity level


```{r setup, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages
library(tidyverse)
library(countrycode)
library(ggplot2)

source("Functions.R")

datadir <- "/Volumes/jgephart/BFA Environment 2/Data"

#_______________________________________________________________________________________________________________________#
# Load lca data
#_______________________________________________________________________________________________________________________#
lca <- clean.lca(LCA_data = read.csv(file.path(datadir, "LCA_compiled_20201109.csv"), stringsAsFactors = FALSE))

#_______________________________________________________________________________________________________________________#
# Load feed FP data 
#_______________________________________________________________________________________________________________________#
# Units of footprint are per 1 kg (apart from energy, which has per L) FIX IT: Rob, is this correct?
feed_fp <- read.csv(file.path(datadir, "Feed_FP_raw.csv"), stringsAsFactors = FALSE)
# Tidy FP data
feed_fp <- clean.feedFP(feed_fp)

#_______________________________________________________________________________________________________________________#
# Load data for on-farm N and P emissions calcs
#_______________________________________________________________________________________________________________________#
# Calculate C,N,P content of fish (based on energy and fat content) and compute discharge based on Czamanski et al 2011, Marine Biology,
#Carbon, nitrogen and phosphorus elemental stoichiometry in aquacultured and wild-caught fish and consequences for pelagic nutrient dynamics and also a simplified NPZ method approach

# N and P content of fish in % of DM
fishNutrition <- read.csv(file.path(datadir, "FishGenus.csv"), stringsAsFactors = FALSE)
fishNutrition <-fishNutrition %>%filter(Preparation=='raw')   #filter out all non raw observations (e.g. dried, cooked)

fishNutrition <- fishNutrition %>%
  mutate(
    N = fishN(Water,Energy.total.metabolizable.calculated.from.the.energy.producing.food.components.original.as.from.source.kcal),
    P = fishP(Water,Energy.total.metabolizable.calculated.from.the.energy.producing.food.components.original.as.from.source.kcal),
    N_fat = fishN_viaFat(Water,Fat.total),
    P_fat = fishP_viaFat(Water,Fat.total),
    N_built_in = Nitrogen.total*100/(100-Water),
    P_built_in = Phosphorus/1000*100/(100-Water)) %>% #conversion to units of percentages. Nitrogen.total is in units of g/100 edible gram; Phosphorous in the dataset is in mg/100 edible gram
  group_by(Scientific.Name) %>%
  summarise(N = mean(N, na.rm = TRUE), P = mean(P, na.rm = TRUE), 
            N_fat = mean(N_fat, na.rm = TRUE), P_fat = mean(P_fat, na.rm = TRUE), 
            N_built_in = mean(N_built_in, na.rm = TRUE), .groups = "drop") 

lca <- lca %>%
  left_join(fishNutrition, by = "Scientific.Name")

#N and P content of feed in % of DM (read from feed composition tables, USDA 1982)
feedNutrition <-clean_feedNutrition(feedNutrition_data = 
                                      read.csv(file.path(datadir, "United-States-Canadian-Tables-of-Feed-1982-pages-68-921-with_CrudeProtein.csv"),
                                               stringsAsFactors = FALSE))

#_______________________________________________________________________________________________________________________#
# Load pond area and yield data
#_______________________________________________________________________________________________________________________#
# Add random pond area as placeholder
lca$harvest <- runif(nrow(lca), min = 10, max = 500)

#_______________________________________________________________________________________________________________________#
# Load on-farm energy data
#_______________________________________________________________________________________________________________________#
energy_gwp <- read.csv(file.path(datadir, "electricity_GWP.csv"))
# Clean eneergy data
energy_gwp$iso3c <- countrycode(energy_gwp$Country, origin = "country.name", destination = "iso3c")

#_______________________________________________________________________________________________________________________#
# Load and clean evaporation data (need to check units and add variability)
#_______________________________________________________________________________________________________________________#
evap <- read.csv("Data/clim_summarise_by_country.csv")
evap$iso3c <- countrycode(evap$admin, origin = "country.name", destination = "iso3c")
evap <- evap %>%
  rename(evap_rate = country_level_mean) %>%
  select(iso3c, evap_rate) %>%
  filter(!is.na(iso3c))

```

## Calculate Feed Footprint

The feed footprint estimate inputs are:

* FCR
* Percent soy, other crops, FMFO and animal
* The GHG, water, N, P, and land footprint per unit of feed ingredient

We then calculate the feed-associated footprint ($FP_{feed}$) as:
$$FP_{feed} = FCR_{dry} \sum_{i=1}^4 FP_{i} p_i$$
where $FCR_{dry}$ is the dry weight feed conversion ratio, $i$ indexes the feed ingredient, $FP_i$ represents the footprint of the feed ingredient, and $p_i$ represents the proportion of the feed comprised of component $i$. 

```{r}
lca$feed.ghg <- estimate.feedFP(LCA_data = lca, Feed_data = feed_fp, FP_option = "Carbon", allocation = "Mass/physical")
lca$feed.water <- estimate.feedFP(LCA_data = lca, Feed_data = feed_fp, FP_option = "Water", allocation = "Mass/physical")
lca$feed.N <- estimate.feedFP(LCA_data = lca, Feed_data = feed_fp, FP_option = "Nitrogen", allocation = "Mass/physical")
lca$feed.P <- estimate.feedFP(LCA_data = lca, Feed_data = feed_fp, FP_option = "Phosphorus", allocation = "Mass/physical")
lca$feed.land <- estimate.feedFP(LCA_data = lca, Feed_data = feed_fp, FP_option = "Land", allocation = "Mass/physical")
```

For species without LCA data, we will need to estimate FCR and the percent soy, other crops, FMFO and animal products in feeds. 

## Calculate on farm footprints

The inputs are: 

* N and P content of protein [treated as constants]
* Protein content of each feed ingredient [treated as constants]
* Protein content of fish [treated as constants]
* Country-specific GHGs with electricity use [treated as constants]
* Diesel, petrol, and natural gas GHG values [treated as constants]
* Yield
* Total harvest
* Production system type
* Aerated or not
* Electricity, diesel, petrol, and natural gas use
* Grow-out period


### Nitrogen and Phosphorus

Alon - update this section to describe the methods from your calcs (in a narrative style) so we have them for the methods section of the paper

The non-feed (which here we mean as the virtual footprint associated with the feed) nitrogen and phosphorus are calculated as by estimating the difference between the N and P in the feeds and the N and P in the final fish, following:
$$FP_{nonfeedN} = FCR_{dry} \sum_{i=1}^4 (N_{i} p_i) - N_{fish}$$
where $N_{Pr}$ represents the average nitrogen content of protein, $Pr_i$ represents the protein content of each feed component, and $Pr_{fish}$ represents the protein content of a unit of fish or shellfish. Similarly, 

$$FP_{nonfeedP} = FCR_{dry} \sum_{i=1}^4 (P_{i} p_i) - P_{fish}$$
where $P_{Pr}$ represents the average phosphorus content of protein.

### Greenhouse gases

The non-feed associated greenhouse gas emissions are calculated as the electricity use times the country-specific GHG footprint, plus the diesel, petrol, and natural gas use times each of their GHG footprint factors. 

### Land

The non-feed associated land use refers to the pond area allocated to the growth of a unit of output. This is calculated as: 
$$FP_{nonfeedland} = Yield/Harvest$$

### Water 

To calculate the on farm water use, we estimate the evaporative losses over the surface area allocated to the unit of production as:
$$FP_{nonfeedwater} = Evap_{rate}FP_{nonfeedland}GrowOut$$

[Sorry I got lazy with equation notation -- will eventually improve!]

```{r}
# Estimate N and P content of feed
lca$feed.N.percent <- estimate.feed.NP(LCA_data = lca, Feed_data = feedNutrition, FP = "N")
lca$feed.P.percent <- estimate.feed.NP(LCA_data = lca, Feed_data = feedNutrition, FP = "P")

# Apply model to estimate N and P lost to the environment
# N and P release in kg 
lca <- estimate.onfarm.NP(LCA_data = lca)

# Estimate on farm GHG
lca$onfarm.ghg <- estimate.onfarm.GHG(LCA_data = lca, energy_data = energy_gwp)

# Estimate on farm land
lca$onfarm.land <- estimate.onfarm.land(LCA_data = lca)

# Estimate on farm water
lca$onfarm.water <- estimate.onfarm.water(LCA_data = lca, evap_data = evap)

```


## Plots

This is largely fake data, so probably shouldn't think about it too much.

```{r}
fp_summary <- lca %>%
  select(Scientific.Name, starts_with("feed."), starts_with("onfarm.")) %>%
  pivot_longer(feed.ghg:onfarm.water, names_to = "FP.component", values_to = "FP.value") %>%
  separate(FP.component, into = c("FP.component", "FP")) %>%
  filter(FP.value > 0)


ggplot(fp_summary, aes(x = Scientific.Name, y = FP.value, fill = FP.component)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~FP, scales = "free_x")
```


